---
title: "Summarize Bounds Tables"
output:
  rmarkdown::html_vignette:
    css: "custom.css"
    df_print: paged
    toc: true
    toc_depth: 2
bibliography: gsDesign.bib
vignette: |
  %\VignetteIndexEntry{Summarize Bounds Tables}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tibble)
library(gt)
library(gsdmvn)
```

This vignette introduce the tutorial to summarize the bounds table in the group sequential design by using `gsdmvn`.  
The function used to generate bounds tables is `gsdmvn::summary_bound()`. 
Then users can use `gsdmvn::as_gt()` to format the above R table into a gt table. 

In this vignette, we introduce how to summarize the bounds table by

- average hazard ratio (AHR) method, where `gsdmvn::gs_design_ahr()` is used;
- weighted logrank (WLR) method, where `gsdmvn::gs_design_wlr()` is used;
- max combo (COMBO) method, where `gsdmvn::gs_design_combo()` is used.

The design parameters we use in this vignette is
```{r}
# enrollment/failure rates
enrollRates <- tibble::tibble(Stratum = "All", 
                      duration = 12, 
                      rate = 1)
failRates <- tibble::tibble(Stratum = "All", duration = c(4, 100), 
                    failRate = log(2) / 12,
                    hr = c(1, .6), 
                    dropoutRate = .001)
# Information fraction
IF <- (1:3)/3 
# Analysis times in months; first 2 will be ignored as IF will not be achieved
analysisTimes <- c(.01, .02, 36)  
# Experimental / Control randomization ratio
ratio <- 1 
# 1-sided Type I error
alpha <- 0.025 
# Type II error (1 - power)
beta <- .1 
# Upper bound
upper <- gsdmvn::gs_spending_bound
upar <- list(sf = gsDesign::sfLDOF, total_spend = 0.025, param = NULL, timing = NULL)
# Lower bound
lower <- gsdmvn::gs_spending_bound
lpar <- list(sf = gsDesign::sfHSD, total_spend = 0.1, param = 0, timing = NULL)
# weight function in WLR
wgt00 <- function(x, arm0, arm1){
                 gsdmvn:::wlr_weight_fh(x, arm0, arm1, rho = 0, gamma = 0)}
wgt05 <- function(x, arm0, arm1){
                 gsdmvn:::wlr_weight_fh(x, arm0, arm1, rho = 0, gamma = .5)}
# test in COMBO
fh_test <- rbind(
      data.frame(rho = 0, gamma = 0, tau = -1, test = 1, Analysis = 1:3, analysisTimes = c(12, 24, 36)), 
      data.frame(rho = c(0, 0.5), gamma = 0.5, tau = -1, test = 2:3, Analysis = 3, analysisTimes = 36)
      )
```

# AHR method 

By using the design parameters above, one can generate an AHR model by `gs_design_ahr` as
```{r}
x_ahr <- gs_design_ahr(
   enrollRates = enrollRates,
   failRates = failRates,
   IF = IF, 
   analysisTimes = analysisTimes, 
   ratio = ratio, 
   alpha = alpha, 
   beta = beta, 
   upper = upper,
   upar = upar,
   lower = lower,
   lpar = lpar
)
```

The outputs of `x_ahr` is a list including

- `enrollRates`: the enrollment rates
- `failRates`: the failure rates
- `bounds`: the summary of the crossing probabilities under H0 \& H1
- `analysis`: the summary of each analysis

The direct output of `x_ahr$bounds` is
```{r}
x_ahr$bounds %>% gt::gt()
```

The direct output of `x_ahr$analysis` is 
```{r}
x_ahr$analysis %>% gt::gt()
```

To summarize the bounds table of `x_ahr`, one can use `summary_bound()`. 
The default output of `summary_bound()` for the AHR method is
```{r}
summary_bound(x_ahr) %>% gt::gt()
```

In the above default table, the variables used to summarize each analysis includes `Analysis`, `Time`, `N`(sample size), `Events`, `AHR`, and `IF` (information fraction). 
But users can custom these variables (and their decimals) displayed as the analysis summary by using the arguments `analysis_vars = ...` and `analysis_decimals = ...`. 
For example
```{r}
summary_bound(
   x_ahr,
   analysis_vars = c("N", "Events"),
   analysis_decimals = c(1, 1)) %>% 
   gt::gt()
```
Please note that there is no need to input `"Analysis"` into `analysis_vars = ...`.

Besides, users can also custom the bound names. 
In the default output, the bound name is `c("Efficacy", "Futility")`, which can be changed into 
`c("A is better", "B is better")` for a 2-sided design by using the argument `bound_names = ...`.
For example,
```{r}
summary_bound(
   x_ahr,
   bound_names = c("A is better", "B is better")) %>% 
   gt::gt()
```

Finally, users can also use `as_gt()` to get the the above R table into a gt table.
Furthermore, they can edit the title/subtitle/spanner/footnotes of the gt table by using the arguments in `summary_bound`. 
```{r}
summary_bound(x_ahr) %>% 
   as_gt(title = "Summary of the Crossing Probability",
         subtitle = "by Using gs_design_ahr",
         colname_spanner = "Cumulative boundary crossing probability",
         colname_spannersub = c("Alternate hypothesis", "Null hypothesis"),
         footnote = list(content = c("approximate hazard ratio to cross bound.", 
                                     "gs_design_ahr is a function in gsdmvn.",
                                     "AHR is average hazard ratio; IF is information fraction."),
                         location = c("~HR at bound", NA, NA),
                         attr = c("colname", "subtitle", "analysis")))
```

The above objective can also be realized by using functions in the R package `gt` for custom design of table layout. 
In the future, if we want to develop rtf, then we can use `as_rtf()`.

# WLR method 

In this section, we discuss the procedure to generate a bound table under the frame work of WLR method.

```{r}
x_wlr <- gs_design_wlr(
   enrollRates = enrollRates,
   failRates = failRates,
   weight = wgt05, 
   # Information fraction
   IF = NULL, 
   # Analysis times in months; first 2 will be ignored as IF will not be achieved
   analysisTimes = sort(unique(x_ahr$analysis$Time)), 
   # Experimental / Control randomization ratio
   ratio = ratio, 
   # 1-sided Type I error
   alpha = alpha, 
   # Type II error (1 - power)
   beta = beta,   
   upper = upper,
   upar = upar,
   lower = lower,
   lpar = lpar
)
```

One can use `x_wlr$bounds`/`x_wlr$analysis` to get a simplified bounds/analysis table, which is similar to that by `x_ahr`.
To get a comprehensive bounds summary table, one can use 
```{r}
summary_bound(x_wlr) %>% gt::gt()
```

The above table can also be format by `as_gt()` as
```{r}
summary_bound(x_wlr) %>% as_gt()
```

# MAX COMBO method 

In this section, we discuss the procedure to generate a bound table under the frame work of MAX COMBO method.

```{r}
x_combo <- gsdmvn::gs_design_combo(
  ratio = 1, 
  alpha = 0.025, 
  beta = 0.2, 
  enrollRates = tibble::tibble(Stratum = "All", duration = 12, rate = 500/12),
  failRates = tibble::tibble(Stratum = "All", duration = c(4, 100), 
                             failRate = log(2) / 15, hr = c(1, .6), dropoutRate = .001), 
  fh_test = fh_test,
  upper = gsdmvn::gs_spending_combo,
  upar = list(sf = gsDesign::sfLDOF, total_spend = 0.025),
  lower = gsdmvn::gs_spending_combo,
  lpar = list(sf = gsDesign::sfLDOF, total_spend = 0.2))
```

One can use `x_combo$enrollRates`, `x_combo$failRates`, `x_combo$bounds`/`x_combo$analysis` to get a simplified enrollment rates/failure rates/bounds/analysis table, which is similar to that by `x_ahr`.
To get a comprehensive bounds summary table, one can use 

```{r}
summary_bound(x_combo) %>% gt::gt()
```

One can also use `as_gt()` to custom the gt style of the bounds table.
```{r}
summary_bound(x_combo) %>% as_gt()
```

