---
title: "Trial design with spending under NPH"
output:
  html_vignette
vignette: >
  %\VignetteIndexEntry{Trial design with spending under NPH}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

This vignette covers how to implement designs for trials with spending assuming non-proportional hazards.
We are primarily concerned with practical issues of implementation rather than design strategies, but we will not ignore design strategy.

## Scenario for consideration

Here we set up enrollment, failure and dropout rates along with assumptions for enrollment duration and times of analyses.

```{r, message=FALSE, warning=FALSE}
library(gsdmvn)
library(dplyr)
library(tibble)
library(gsDesign)
library(gsDesign2)
library(gt)
```

```{r}
analysisTimes <- c(18, 24, 30, 36)
enrollRates <- tibble::tibble(
  Stratum = "All",
  duration = c(2, 2, 2, 6),
  rate = c(8, 12, 16, 24)
)
failRates <- tibble::tibble(
  Stratum = "All",
  duration = c(3, 100),
  failRate = log(2) / c(8, 14),
  hr = c(.9, .6),
  dropoutRate = .001
)
```

## Deriving power for a given sample size

We derive statistical information at targeted analysis times.

```{r}
xx <- gsDesign2::AHR(enrollRates = enrollRates, failRates = failRates, totalDuration = analysisTimes)
Events <- ceiling(xx$Events)
yy <- gs_info_ahr(enrollRates = enrollRates, failRates = failRates, events = Events)
```

Now we can examine power using `gs_power_npe()`:

```{r}
timing <- yy$info0 / max(yy$info0)
d <- gsDesign::gsDesign(k = length(timing), test.type = 2, sfu = sfLDOF, alpha = .025, timing = timing)
zz <- gs_power_npe(
  theta = yy$theta, 
  info = yy$info, info0 = yy$info0,
  upper = gs_b, lower = gs_b,
  upar = d$upper$bound,
  lpar = d$lower$bound
)
zz %>% 
  gt() %>% 
  fmt_number(columns = 3:10, decimals = 4)
```

## Deriving sample size to power a trial

If we were using a fixed design, we would approximate the sample size as follows:

```{r}
K <- 4
minx <- ((qnorm(.025) / sqrt(zz$info0[K]) + qnorm(.1) / sqrt(zz$info[K])) / zz$theta[K])^2
minx
```

If we inflate the enrollment rates by `minx` and use a fixed design, we will see this achieves the targeted power.

```{r}
gs_power_npe(
  theta = yy$theta[K], info = yy$info[K] * minx, info0 = yy$info0[K] * minx,
  upar = qnorm(.975), lpar = -Inf) %>%
  subset(hypothesis == "H1") %>% 
  filter(Bound == "Upper") %>%
  gt() %>% 
  fmt_number(columns = 3:10, decimals = 4)
```

The power for a group sequential design with the same final sample size is a bit lower:

```{r}
zz <- gs_power_npe(
  theta = yy$theta, info = yy$info * minx, info0 = yy$info0 * minx,
  upper = gs_spending_bound, lower = gs_spending_bound,
  upar = list(sf = gsDesign::sfLDOF, total_spend = 0.025, param = NULL, timing = NULL),
  lpar = list(sf = gsDesign::sfLDOF, total_spend = 0.025, param = NULL, timing = NULL)
)
zz %>% 
  subset(hypothesis == "H1") %>% 
  select(-hypothesis) %>% 
  gt() %>% 
  fmt_number(columns = 3:10, decimals = 4)
```

If we inflate this a bit we will be overpowered.

```{r}
zz <- gs_power_npe(
  theta = yy$theta, 
  info = yy$info * minx * 1.2, 
  info0 = yy$info0 * minx * 1.2,
  upper = gs_spending_bound, 
  lower = gs_spending_bound,
  upar = list(sf = gsDesign::sfLDOF, total_spend = 0.025, param = NULL, timing = NULL),
  lpar = list(sf = gsDesign::sfLDOF, total_spend = 0.025, param = NULL, timing = NULL)
)
zz %>% 
  subset(hypothesis == "H1") %>% 
  select(-hypothesis) %>% 
  gt() %>% 
  fmt_number(columns = 3:10, decimals = 4)
```

Now we use `gs_design_npe()` to inflate the information proportionately to power the trial.

```{r}
theta <- yy$theta
info <- yy$info
info0 <- yy$info0
upper <- gs_spending_bound
lower <- gs_spending_bound
upar <- list(sf = gsDesign::sfLDOF, total_spend = 0.025, param = NULL, timing = NULL)
lpar <- list(sf = gsDesign::sfLDOF, total_spend = 0.025, param = NULL, timing = NULL)
alpha <- .025
beta <- .1
binding <- FALSE
test_upper <- TRUE
test_lower <- TRUE
r <- 18
tol <- 1e-06

zz <- gs_design_npe(
  theta = yy$theta, info = yy$info, info0 = yy$info0,
  upper = gs_spending_bound, lower = gs_spending_bound,
  upar = list(sf = gsDesign::sfLDOF, total_spend = 0.025, param = NULL, timing = NULL),
  lpar = list(sf = gsDesign::sfLDOF, total_spend = 0.025, param = NULL, timing = NULL)
)
zz %>% 
  subset(hypothesis == "H1") %>% 
  select(-hypothesis) %>% 
  gt() %>% 
  fmt_number(columns = 3:10, decimals = 4)
```
