---
title: "Fixed Design"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 4
    number_sections: yes
bibliography: gsDesign.bib
vignette: |
  %\VignetteIndexEntry{Group Sequential Bound Tables}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---


```{r, message=FALSE, warning=FALSE}
library(gsDesign)
library(tibble)
library(gt)
# load the develop version of gsdmvn
# it will finally be replaced by `libraray(gsdmvn)`
devtools::load_all()
```

# Parameters
```{r}
# Enrollment rate
enrollRates <- tibble::tibble(
   Stratum = "All", 
   duration = 18, 
   rate = 20)

# Failure rates
failRates <- tibble::tibble(
   Stratum = "All", 
   duration = c(4, 100), 
   failRate = log(2) / 12,
   hr = c(1, .6), 
   dropoutRate = .001)

# Study duration in months
studyDuration <- 36

# Experimental / Control randomization ratio
ratio <- 1 

# 1-sided Type I error
alpha <- 0.025 
# Type II error (1 - power)
beta <- 0.1 
```

# AHR {.tabset}

## under fixed power 
```{r}
x <- fixed_design(x = "AHR", 
             alpha = alpha, power = 1 - beta, 
             enrollRates = enrollRates, failRates = failRates, 
             studyDuration = studyDuration, ratio = ratio) 

x %>% summary()

x %>% summary() %>% as_gt()
```

## under fixed sample size
```{r}
fixed_design(x = "AHR", 
             alpha = alpha,  
             enrollRates = enrollRates, failRates = failRates, 
             studyDuration = studyDuration, ratio = ratio) %>% 
  summary() %>% 
  as_gt()
```

# FH {.tabset}

## under fixed power (default rho/gamma)
```{r}
# fixed design with a given power with default rho/gamma
x <- fixed_design(x = "FH", 
             alpha = alpha, power = 1 - beta, 
             enrollRates = enrollRates, failRates = failRates, 
             studyDuration = studyDuration, ratio = ratio) 
x %>% summary() 
x %>% summary() %>% as_gt()
```

## under fixed power (custom rho/gamma)
```{r}
# fixed design with a given power with input rho/gamma
fixed_design(x = "FH", 
             alpha = alpha, power = 1 - beta, 
             enrollRates = enrollRates, failRates = failRates, 
             studyDuration = studyDuration, ratio = ratio,
             rho = 0.5, gamma = 0.5) %>% 
  summary() %>%  
  as_gt()
```


## under fixed sample size (default rho/gamma)
```{r}
# fixed design with power calculated
fixed_design(x = "FH", 
             alpha = alpha,  
             enrollRates = enrollRates, failRates = failRates, 
             studyDuration = studyDuration, ratio = ratio) %>% 
  summary() %>%  
  as_gt()
```

## under fixed sample size (custom rho/gamma)
```{r}
# fixed design with power calculated
fixed_design(x = "FH", 
             alpha = alpha,  
             enrollRates = enrollRates, failRates = failRates, 
             studyDuration = studyDuration, ratio = ratio,
             rho = 0.5, gamma = 0.5) %>% 
  summary() %>%  
  as_gt()
```

# MB {.tabset}

## under fixed power (default rho/gamma/tau)
```{r}
x <- fixed_design(x = "MB", 
             ratio = ratio, 
             alpha = alpha, power = 1 - beta,
             enrollRates = enrollRates,
             failRates = failRates,
             studyDuration = studyDuration) 
x %>% summary() 
x %>% summary() %>% as_gt()
```

## under fixed power (custom rho/gamma/tau)
```{r}
fixed_design(x = "MB", 
             ratio = ratio, 
             alpha = alpha, power = 1 - beta,
             enrollRates = enrollRates,
             failRates = failRates,
             studyDuration = studyDuration,
             rho = 0.5, gamma = 0.5, tau = 4) %>% 
  summary() %>%  
  as_gt()
```


## under fixed sample size (default rho/gamma/tau)
```{r}
fixed_design(x = "MB",
             ratio = ratio,
             alpha = alpha,
             enrollRates = enrollRates,
             failRates = failRates,
             studyDuration = studyDuration) %>% 
  summary() %>%  
  as_gt()
```

## under fixed sample size (custom rho/gamma/tau)
```{r}
fixed_design(x = "MB",
             ratio = ratio,
             alpha = alpha,
             enrollRates = enrollRates,
             failRates = failRates,
             studyDuration = studyDuration,
             rho = 0.5, gamma = 0.5, tau = 4) %>% 
  summary() %>%  
  as_gt()
```

# LF {.tabset}

## under fixed power
```{r}
fixed_design(x = "LF", alpha = alpha, power = 1 - beta, 
             ratio = ratio,
             enrollRates = enrollRates,
             failRates = failRates,
             studyDuration = studyDuration) %>% 
  summary() %>%  
  as_gt()
```


## under sample size
```{r}
fixed_design(x = "LF", alpha = alpha,  
             ratio = ratio, 
             enrollRates = enrollRates,
             failRates = failRates,
             studyDuration = studyDuration) %>% 
  summary() %>%  
  as_gt()
```

# MaxCombo {.tabset}

## under fixed power (default rho/gamma/tau)
```{r}
x <- fixed_design(x = "MaxCombo", alpha = alpha, power = 1 - beta,
             ratio = ratio, 
             enrollRates = enrollRates,
             failRates = failRates,
             studyDuration = studyDuration) 
x %>% summary() 
x %>% summary() %>% as_gt()
```

## under fixed power (custom rho/gamma/tau)
```{r}
fixed_design(x = "MaxCombo", alpha = alpha, power = 1 - beta,
             ratio = ratio, 
             enrollRates = enrollRates,
             failRates = failRates,
             studyDuration = studyDuration,
             rho = c(0, 0.5, 0.5),
             gamma = c(0, 0, 0.5),
             tau = c(-1, 4, 6)) %>% 
  summary() %>% 
  as_gt()
```


## under sample size (default rho/gamma/tau)

```{r}
fixed_design(x = "MaxCombo", alpha = alpha, 
             ratio = ratio, 
             enrollRates = enrollRates,
             failRates = failRates,
             studyDuration = studyDuration) %>% 
  summary() %>% 
  as_gt()
```


## under sample size (custom rho/gamma/tau)
```{r}
fixed_design(x = "MaxCombo", alpha = alpha, 
             ratio = ratio, 
             enrollRates = enrollRates,
             failRates = failRates, 
             studyDuration = studyDuration,
             rho = c(0, 0.5, 0.5),
             gamma = c(0, 0, 0.5),
             tau = c(-1, 4, 6)) %>% 
  summary() %>% 
  as_gt()
```


# Multiple Designs

```{r, message = FALSE}
x_AHR <- fixed_design(x = "AHR", alpha = alpha, ratio = ratio,
                      enrollRates = enrollRates, failRates = failRates, 
                      studyDuration = studyDuration)

x_FH <- fixed_design(x = "FH", alpha = alpha, ratio = ratio, 
                     enrollRates = enrollRates, failRates = failRates, studyDuration = studyDuration,
                     rho = 0.5, gamma = 0.5) 

x_MB <- fixed_design(x = "MB", alpha = alpha, ratio = ratio, 
                     enrollRates = enrollRates,failRates = failRates, studyDuration = studyDuration, 
                     rho = 0.5, gamma = 0.5, tau = 4)

x_LF <- fixed_design(x = "LF", alpha = alpha, ratio = ratio, 
                     enrollRates = enrollRates, failRates = failRates, studyDuration = studyDuration)

x_MaxCombo <- fixed_design(x = "MaxCombo", alpha = alpha, ratio = ratio, 
                           enrollRates = enrollRates, failRates = failRates, studyDuration = studyDuration,
                           rho = c(0, 0.5, 0.5), gamma = c(0, 0, 0.5), tau = c(-1, 4, 6))

rbind(summary(x_AHR), summary(x_FH), summary(x_MB), summary(x_LF), summary(x_MaxCombo)) %>% gt()
```

