---
title: "`info_scale` in `gs_power_npe`"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    toc: yes
    toc_depth: 4
    number_sections: yes
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
library(gsDesign)
library(dplyr)
library(gt)
library(testthat)
devtools::load_all()
```
```{r}
tibble::tibble(argument =         c("theta", "theta1", "info", "info0", "info1"),
               `efficacy bound` = c("-",     "-",      "-",    "-",     "-"),
               `futility bound` = c("-",     "Yes",    "-",    "-",     "Yes"),
               power =            c("Yes",   "-",      "-",    "-",     "-"),
               type1_error =      c("-",     "-",      "-",    "-",     "-"),
               IF =               c("-",     "-",      "-",    "-",     "-")) %>% gt()

tibble::tibble(`if` = c("info_scale = 0", "info_scale = 1", "info_scale = 2"),
               `then` = c("info = info0", "info = info1",   "info = info"),
               `calculation of IF` = c("info0/max(info0)", "info1/max(info1)", "info/max(info)") ) %>% gt()

```

- No matter what `info_scale` is, the lower and upper bound are decide by `info1` and `info0` under both H1 and H0, respectively.
- The `info_scale` will affect 
  + the power by `gs_power_ahr`, `gs_power_wlr`, `gs_power_combo`

# Application to AHR {.tabset}
```{r}
hr <- .6
alpha <- 0.0125
beta <- .1
ratio <- 1
analysisTimes <- 1:2
K <- 2
n_per_arm <- 140
enroll_duration <- .0001
median_tte <- 0.75
enrollRates <- tibble::tibble(Stratum = "All", duration = enroll_duration, rate = 2 * n_per_arm / enroll_duration)
failRates <- tibble::tibble(Stratum = "All",
                            duration = 1,
                            failRate = log(2) / median_tte,
                            hr = .6,
                            dropoutRate = 0)
upper <- gs_spending_bound
upar <- list(sf = gsDesign::sfLDPocock, total_spend = alpha, param = NULL, timing = NULL, max_info = NULL)
lower <- gs_b
lpar <- rep(-Inf, 2)

x <- gs_info_ahr(
    enrollRates = enrollRates,
    failRates = failRates,
    ratio = ratio,
    events = NULL,
    analysisTimes = analysisTimes)
```

## `info_scale = 0`
```{r}
gs_power_npe(
    theta = x$theta, 
    info = x$info, 
    info0 = x$info0,
    info_scale = 0,
    binding = FALSE,
    upper = upper, 
    upar = upar,
    test_upper = TRUE,
    lower = lower, 
    lpar = lpar,
    test_lower = TRUE,
    r = 18, 
    tol = 1e-6) %>% gt() %>% tab_header("info_scale = 0")
```

## `info_scale = 1`
```{r}
gs_power_npe(
    theta = x$theta, 
    info = x$info, 
    info0 = x$info0,
    info_scale = 1,
    binding = FALSE,
    upper = upper, 
    upar = upar,
    test_upper = TRUE,
    lower = lower, 
    lpar = lpar,
    test_lower = TRUE,
    r = 18, 
    tol = 1e-6) %>% gt() %>% tab_header("info_scale = 1")
```

## `info_scale = 2`
```{r}
gs_power_npe(
    theta = x$theta, 
    info = x$info, 
    info0 = x$info0,
    info_scale = 2,
    binding = FALSE,
    upper = upper, 
    upar = upar,
    test_upper = TRUE,
    lower = lower, 
    lpar = lpar,
    test_lower = TRUE,
    r = 18, 
    tol = 1e-6) %>% gt() %>% tab_header("info_scale = 2")
```



# Implementation of `gs_power_npe`

## Default {.tabset}

The default is a single analysis with type I error controlled:

- `theta` = `theta1` = 0.1
- `info` = `info1` = `info0` = 1
- upper bound is `qnorm(.975)`
- lower bound is `-Inf` and it is non-binding.

In this case, since `info` = `info1` = `info0`, so the above 3 outputs are the same.

### info_scale = 2
```{r}
gs_power_npe(info_scale = 2) %>% gt() %>% tab_header(title = "info_scale = 2")
```

### info_scale = 1
```{r}
gs_power_npe(info_scale = 1) %>% gt() %>% tab_header(title = "info_scale = 1")
```

### info_scale = 0
```{r}
gs_power_npe(info_scale = 0) %>% gt() %>% tab_header(title = "info_scale = 0")
```

## Fixed Bound {.tabset}

In the following example, we investigate a design with 3 analysis. 
The upper boundary is fixed and calculated from `gsDesign`.
The lower boundary is `(-1, 0, 0)`.

### info_scale = 2
```{r}
x <- gs_power_npe(
  theta = c(.1, .2, .3),
  info = (1:3) * 40, 
  info0 = (1:3) * 30,
  info1 = (1:3) * 50,
  info_scale = 2,
  upper = gs_b,
  upar = gsDesign::gsDesign(k = 3, sfu = gsDesign::sfLDOF)$upper$bound,
  lower = gs_b, 
  lpar = c(-1, 0, 0)) 

x %>% subset(Bound == "Upper") %>% gt() %>% tab_header(title = "upper bound under info_scale = 2")
x %>% subset(Bound == "Lower") %>% gt() %>% tab_header(title = "lower bound under info_scale = 2")
```

### info_scale = 1
```{r}
x <- gs_power_npe(
  theta = c(.1, .2, .3),
  info = (1:3) * 40, 
  info0 = (1:3) * 30,
  info1 = (1:3) * 50,
  info_scale = 1,
  upper = gs_b,
  upar = gsDesign::gsDesign(k = 3, sfu = gsDesign::sfLDOF)$upper$bound,
  lower = gs_b, 
  lpar = c(-1, 0, 0)) 

x %>% subset(Bound == "Upper") %>% gt() %>% tab_header(title = "upper bound under info_scale = 1")
x %>% subset(Bound == "Lower") %>% gt() %>% tab_header(title = "lower bound under info_scale = 1")
```

### info_scale = 0
```{r}
x <- gs_power_npe(
  theta = c(.1, .2, .3),
  info = (1:3) * 40, 
  info0 = (1:3) * 30,
  info1 = (1:3) * 50,
  info_scale = 0,
  upper = gs_b,
  upar = gsDesign::gsDesign(k = 3, sfu = gsDesign::sfLDOF)$upper$bound,
  lower = gs_b, 
  lpar = c(-1, 0, 0))

x %>% subset(Bound == "Upper") %>% gt() %>% tab_header(title = "upper bound under info_scale = 0")
x %>% subset(Bound == "Lower") %>% gt() %>% tab_header(title = "lower bound under info_scale = 0")
```

## Spending Function Bounds {.tabset}

### info_scale = 2 
```{r}
x <- gs_power_npe(
  theta = c(.1, .2, .3),
  info = (1:3) * 40,
  info0 = (1:3) * 30,
  info1 = (1:3) * 50,
  info_scale = 2,
  upper = gs_spending_bound,
  upar = list(sf = gsDesign::sfLDOF, total_spend = 0.025, param = NULL, timing = NULL),
  lower = gs_spending_bound,
  lpar = list(sf = gsDesign::sfHSD, total_spend = 0.1, param = -1, timing = NULL)) 

x %>% subset(Bound == "Upper") %>% gt() %>% tab_header(title = "upper bound under info_scale = 2")
x %>% subset(Bound == "Lower") %>% gt() %>% tab_header(title = "lower bound under info_scale = 2")
```

### info_scale = 1
```{r}
x <- gs_power_npe(
  theta = c(.1, .2, .3),
  info = (1:3) * 40,
  info0 = (1:3) * 30,
  info1 = (1:3) * 50,
  info_scale = 1,
  upper = gs_spending_bound,
  upar = list(sf = gsDesign::sfLDOF, total_spend = 0.025, param = NULL, timing = NULL),
  lower = gs_spending_bound,
  lpar = list(sf = gsDesign::sfHSD, total_spend = 0.1, param = -1, timing = NULL)) 

x %>% subset(Bound == "Upper") %>% gt() %>% tab_header(title = "upper bound under info_scale = 1")
x %>% subset(Bound == "Lower") %>% gt() %>% tab_header(title = "lower bound under info_scale = 1")
```

### info_scale = 0
```{r}
x <- gs_power_npe(
  theta = c(.1, .2, .3),
  info = (1:3) * 40,
  info0 = (1:3) * 30,
  info1 = (1:3) * 50,
  info_scale = 1,
  upper = gs_spending_bound,
  upar = list(sf = gsDesign::sfLDOF, total_spend = 0.025, param = NULL, timing = NULL),
  lower = gs_spending_bound,
  lpar = list(sf = gsDesign::sfHSD, total_spend = 0.1, param = -1, timing = NULL)) 

x %>% subset(Bound == "Upper") %>% gt() %>% tab_header(title = "upper bound under info_scale = 0")
x %>% subset(Bound == "Lower") %>% gt() %>% tab_header(title = "lower bound under info_scale = 0")
```




# Tests of `gs_power_npe`

## Compare with `mvtnorm`
```{r}
# from gs_power_npe()
info <- c(40, 100)
x <- gs_power_npe(theta = 0,
                  info = info,
                  info0 = NULL,
                  info_scale = 1,
                  binding = FALSE,
                  upper = gs_spending_bound,
                  upar = list(sf = gsDesign::sfLDOF, param = NULL, total_spend = 0.025),
                  lower = gs_spending_bound,
                  lpar = list(sf = gsDesign::sfLDOF, param = NULL, total_spend = 0.02))

x_upper <- x %>% filter(Bound == "Upper", hypothesis == "H1")
x_lower <- x %>% filter(Bound == "Lower", hypothesis == "H1")

# from mvtnorm - lower bound
alpha.t <- 0.025
r <- info[1]/info[2]
b.ia <- gsDesign::sfLDOF(alpha = alpha.t, t = r)
alpha.ia <- b.ia$spend

Pb <- function(alpha.t, alpha.ia, r, b){
  temp = mvtnorm::pmvnorm(lower = c(-Inf, b), 
                          upper = c(qnorm(1 - alpha.ia), Inf),
                          corr = rbind(c(1, sqrt(r)), c(sqrt(r), 1)))
  return(alpha.t - alpha.ia - temp)
}
b <- uniroot(Pb, c(1.96, 4), alpha.t = alpha.t, alpha.ia = alpha.ia, r = r)
pb <- 1- pnorm(b$root)

# from mvtnorm - upper bound
beta.t <- 0.02
a.ia <- gsDesign::sfLDOF(alpha = beta.t, t = r)
beta.ia <- a.ia$spend

Pa <- function(beta.t, beta.ia,  r, a){
  temp <- mvtnorm::pmvnorm(lower = c(-Inf, qnorm(beta.ia)), 
                           upper = c(a, Inf),
                           corr = rbind(c(1, sqrt(r)), c(sqrt(r), 1)))
  return(beta.t -  beta.ia - temp)
  }
a <- uniroot(Pa, c(-4, 1.96), beta.t = beta.t, beta.ia = beta.ia, r = r)
pa <- pnorm(a$root)

# compare  
expect_equal(x_upper$Probability, cumsum(c(b.ia$spend, pb)), tolerance = 1e-3)
expect_equal(x_lower$Probability, cumsum(c(a.ia$spend, pa)), tolerance = 1e-3)
```

## Compare with `gsDesign`
```{r}
x <- gsDesign(
    k = 3, delta = 0.1, test.type = 4, 
    alpha = 0.025, beta = 0.1, timing = c(.5, .75),
    sfu = sfLDOF, sfupar = NULL, sfl = sfHSD, sflpar = -2)

x1 <- tibble::tibble(
    Analysis = rep(1:3,2), Bound = c(rep("Upper", 3), rep("Lower", 3)),
    Z = c(x$upper$bound,x$lower$bound),
    Probability = c(cumsum(x$upper$prob[,2]), cumsum(x$lower$prob[,2])),
    theta = rep(x$delta, 6),
    info = rep(x$n.I, 2)
  ) 
  
x2 <- gs_power_npe(
    theta = .1, info = x$n.I, info_scale = 2,
    upper= gs_spending_bound, upar = list(sf = gsDesign::sfLDOF, param = NULL, total_spend = 0.025),
    lower=  gs_spending_bound, lpar = list(sf = gsDesign::sfHSD, param = -2, total_spend = 0.1))  

# compare upper prob
prob_x1 <- (x1 %>% filter(Bound == "Upper"))$Probability
prob_x2 <- (x2 %>% filter(Bound == "Upper", hypothesis == "H1"))$Probability
expect_equal(prob_x1, prob_x2, tol = 1e-5)

# compare lower prob
prob_x1 <- (x1 %>% filter(Bound == "Lower"))$Probability
prob_x2 <- (x2 %>% filter(Bound == "Lower", hypothesis == "H1"))$Probability
expect_equal(prob_x1, prob_x2, tol = 1e-5)
```
