---
title: "Power Evaluation with Spending bounds"
subtitle: "Binomial Trial Example"
output:
  html_vignette
bibliography: gsDesign.bib
vignette: >
  %\VignetteIndexEntry{Power Evaluation with Spending bounds}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  dev = "png", 
  dev.args= list(type = "cairo")
)
library(gsdmvn)
```


## Overview

This vignette covers how to compute power or Type I error for a design derived with a spending bound. 
We will write this with a general non-constant treatment effect using `gs_design_npe()` to derived the design under one parameter setting and computing power under another setting.
We will use a trial with a binary endpoint to enable a full illustration.


## Scenario for consideration

We consider a scenario largely based on the CAPTURE study @CAPTURE where the primary endpoint was a composite of death, acute myocardial infarction or the need for recurrent percutaneous intervention within 30 days of randomization.
That is, we consider a 2-arm trial with an experimental arm and a control arm.
The primary endpoint for the trial is a binary indicator for each participant if they have a failed outcome. 
For this case, we consider the parameter $\theta = p_1 - p _2$ where $p_1$ denotes the probability that a trial participant in the control group experiences a failure and $p_2$ represents the same probability for a trial participant in the experimental group. 
We will assume $K=3$ analyses after 350, 700, and 1400 patients have been observed with equal randomization between the treatment groups. 
The study was designed with approximately 80% power (Type II error $\beta = 1 - 0.8 = 0.2$) and 2.5% one-sided Type I error ($\alpha = 0.025$) to detect a reduction from a 15% event rate ($p_1 = 0.15$) in the control group to 10% ($p_2 = 0.1$) in the experimental group.
The parameter of interest is $\theta = p_1 - p_2$. We denote the alternate hypothesis as 
$$H_1: \theta = \theta_1= p_1^1 - p_2^1 = 0.15 - 0.10 = 0.05$$ and the null hypothesis
$$H_0: \theta = \theta_0 = 0 = p_1^0 - p_2^0$$ where $p^0_1 = p^0_2= (p_1^1+p_2^1)/2 = 0.125$ 
as laid out in @LachinBook. 



```{r}
library(gsdmvn)
p0 <- 0.15 # assumed failure rate in control group
p1 <- 0.10 # assumed failure rate in experimental group
alpha <- 0.025 # Design Type I error
beta <- 0.2 # Design Type II error for 80% power
```

We note that had we considered a success outcome such as objective response in an oncology study, we would let $p_1$ denote the experimental group and $p_2$ the control group response rate.
Thus, we always set up the notation so the $p_1>p_2$ represents superiority for the experimental group.

## Notation and Statistical Testing

We let $X_{ij}\sim \hbox{Bernoulli}(p_i),$ $j=1,2,\ldots,n_{ik}$ denote independent random variables in the control ($i=1$) and experimental ($i=2$) groups through analysis $k=1,2,\ldots,K$ where $n_{i1} < n_{i2}<\ldots<n_K$.
We let $Y_{ik}=\sum_{j=1}^{n_k}X_{ij}$.
At analysis $k=1,2,\ldots$ and for $j=1,2$ we denote the proportion of failures in group $i$ at analysis $k$

$$ \hat{p}_{ik}=Y_{ik}/n_{ik}.$$
Estimating under null hypothesis $H_0: p_1=p_2\equiv  p_0$ we estimate

$$ \hat{p}_{0k}=\frac{Y_{1k}+ Y_{2k}}{n_{1k}+ n_{2k}}=
\frac{n_{1k}\hat p_{1k} + n_{2k}\hat p_{2k}}{n_{1k} + n_{2k}}. $$
We note

$$\hbox{Var}(\hat p_{ik})=\frac{p_{i}(1-p_i)}{n_{ik}},$$
and its consistent estimatator
$$\widehat{\hbox{Var}}(\hat p_{ik})=\frac{\hat p_{ik}(1-\hat p_{ik})}{n_{ik}},$$
$i=1,2$, $k=1,2,\ldots,K$.
Letting $\hat\theta_k=\hat p_{1k}-\hat p_{2k},$ we also have

$$\sigma^2_k\equiv \hbox{Var}(\hat\theta_i)=\frac{p_1(1-p_1)}{n_{1k}}+\frac{p_2(1-p_2)}{n_{2k}},$$
its consistent estimator 
$$\hat\sigma^2_k=\frac{\hat p_{1k}(1-\hat p_{1k})}{n_{1k}}+\frac{\hat p_{2k}(1-\hat p_{2k})}{n_{2k}},$$
and the corresponding null hypothesis estimator
$$\hat\sigma^2_{0k}=\hat p_{0k}(1-\hat p_{0k})\left(\frac{1}{n_{1k}}+ \frac{1}{n_{2k}}\right),$$
$k=1,2,\ldots,K$. Statistical information for each of these quantities and their corresponding estimators are denoted by

$$\begin{align}
\mathcal{I}_k = &1/\sigma^2_k,\\
\mathcal{\hat I}_k = &1/\hat \sigma^2_k,\\
\mathcal{I}_{0k} =& 1/ \sigma^2_{0k},\\
\mathcal{\hat I}_{0k} =& 1/\hat \sigma^2_{0k},
\end{align}$$
$k=1,2,\ldots,K$. 
Testing, as recommended by @LachinBook, is done with the large sample test with the null hypothesis variance estimate and without continuity correction:
$$Z_k = \hat\theta_k/\hat\sigma_{0k}=\frac{\hat p_{1k} - \hat p_{2k}}{\sqrt{(1/n_{1k}+ 1/n_{2k})\hat p_{0k}(1-\hat p_{0k})} },$$
which is asymptotically Normal(0,1) if $p_1=p_2$ and Normal(0, $\sigma_{0k}^2/\sigma_k^2$) more generally for any $p_1, p_2$, $k=1,2,\ldots,K$.
We note that $\chi^2=Z^2_k$ is the $\chi^2$ test without continuity correction as recommended by @CCmyth. 
Note finally that this extends in a straightforward way the non-inferiority test of @FarringtonManning if the null hypothesis is $\theta = p_1 - p_2 - \delta = 0$ for some non-inferiority margin $\delta > 0$; $\delta < 0$ would correspond to what is referred to as super-superiority @Chan2002, requiring that experimental therapy has been shown to be superior to control by at least a margin $-\delta>0$.

## Power Calculations

We begin with a fixed design to simplify.
We follow the approach of @LachinBook to compute power accounting for the different variance (statistical information) computations under the null hypothesis noted in the previous section.
Noting the asymptotic equivalence

$$Z_k\approx \hat\theta_k/\sigma_{0k}=\frac{\hat p_{1k} - \hat p_{2k}}{\sqrt{(1/n_{1k}+ 1/n_{2k})p_{0}(1- p_0)} }.$$
We denote $n_k=n_{1k}+n_{2k},$ $k=1,2,\ldots, K$ and assume a constant proportion $\xi_i$ randomized to each group $i=1,2.$ Thus, 

$$Z_k\approx \frac{\sqrt{n_k}(\hat p_{1k} - \hat p_{2k})}{\sqrt{(1/\xi_1+ 1/\xi_2)p_{0}(1- p_0)} }.$$

we have the asymptotic distribution

$$Z_k\sim\hbox{Normal}\left(\sqrt{n_k}\frac{p_1 - p_2}{\sqrt{(1/\xi_1+ 1/\xi_2) p_0(1- p_0)} },\sigma^2_{0k}/\sigma^2_{1k}\right).$$
We note that 

$$ \sigma^2_{0k}/\sigma^2_{1k} = \frac{ p_0(1-p_0)\left(1/\xi_1+ 1/\xi_2\right)}{p_1(1-p_1)/\xi_1+p_2(1-p_2)/\xi_2}.$$ 
We also note by definition that $\sigma^2_{0k}/\sigma^2_{1k}=\mathcal I_k/\mathcal I_{0k}.$
Based on an input $p_1, p_2, \xi_1, \xi_2 = 1-\xi_1, n_k$ we will compute $\theta, \mathcal{I}_k, \mathcal{I}_{0k}$, $k=1,2,\ldots,K$.


```{r}
library(tibble)
gs_info_binomial <- function(p1, p2, xi1, n, delta = NULL){
  if (is.null(delta)) delta <- p1 - p2
  # Compute (constant) effect size at each analysis theta
  theta <- rep(p1 - p2, length(n))
  # compute null hypothesis rate, p0
  p0 <- xi1 * p1 + (1 - xi1) * p2
  # compute information based on p1, p2
  info <-  n / (p1 * (1 - p1) / xi1 + p2 * (1 - p2) / (1 - xi1))
  # compute information based on null hypothesis rate of p0
  info0 <- n / (p0 * (1 - p0)*(1 / xi1 + 1 / (1 - xi1)))
  # compute information based on H1 rates of p1star, p2star
  p1star <- p0 + delta * xi1
  p2star <- p0 - delta * (1 - xi1)
  info1 <-  n / (p1star * (1 - p1star) / xi1 + p2star * (1 - p2star) / (1 - xi1))
  return(tibble(Analysis = 1:length(n),
                n = n,
                theta = theta,
                theta1 = rep(delta, length(n)),
                info = info,
                info0 = info0,
                info1 = info1))
}
```


For the CAPTURE trial, we have 


```{r}
h1 <- gs_info_binomial(p1 = .15, p2 = .1, xi1 = .5, n = c(350, 700, 1400))
h1
```

Now we examine information for a smaller assumed treatment difference than the alternative:

```{r}
h <- gs_info_binomial(p1 = .15, p2 = .12, xi1 = .5, delta = .05, n = c(350, 700, 1400))
h
```

We can plug these into `gs_power_npe()` with the intended spending functions.
We begin with power under the alternate hypothesis

```{r}
gs_power_npe(theta = h1$theta, theta1 = h1$theta, info = h1$info,
             info0 = h1$info0, info1 = h1$info1,
             upper = gs_spending_bound,
             upar = list(sf = gsDesign::sfLDOF, total_spend = 0.025),
             lower = gs_spending_bound,
             lpar = list(sf = gsDesign::sfHSD, param = -2, total_spend = 0.2)
)
```

```{r}
gs_power_npe(theta = h$theta, theta1 = h$theta1, info = h$info,
             info0 = h$info0, info1 = h$info1,
             upper = gs_spending_bound,
             upar = list(sf = gsDesign::sfLDOF, total_spend = 0.025),
             lower = gs_spending_bound,
             lpar = list(sf = gsDesign::sfHSD, param = -2, total_spend = 0.2)
)
```

# References
