---
title: "Updated gs_design_xxx and Nice Bounds Summary"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 4
    toc_float: true
    number_sections: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tibble)
library(gt)
devtools::load_all()
```

This vignette discusses the usage of updated 

- `gs_power_npe()` 
- `gs_design_npe()` 
- `gs_design_ahr()`
- `gs_design_wlr()`

The different between the new version and the old version is that, when you set the argument `calc_ho_prob = TRUE`, it calculate the crossing probability under the null hypothesis.

For `gs_design_combo()`, it is not updated, since it already reports the crossing probability under the null hypothesis.

- Can we add IF in `gs_design_combo()`?
- How about theta? (Different test have different theta)

The design parameters we use in this vignette is
```{r}
# enrollment/failure rates
enrollRates <- tibble::tibble(Stratum = "All", 
                      duration = 12, 
                      rate = 1)
failRates <- tibble::tibble(Stratum = "All", duration = c(4, 100), 
                    failRate = log(2) / 12,
                    hr = c(1, .6), 
                    dropoutRate = .001)
# Information fraction
IF <- (1:3)/3 
# Analysis times in months; first 2 will be ignored as IF will not be achieved
analysisTimes <- c(.01, .02, 36)  
# Experimental / Control randomization ratio
ratio <- 1 
# 1-sided Type I error
alpha <- 0.025 
# Type II error (1 - power)
beta <- .1 
# Upper bound
upper <- gsdmvn::gs_spending_bound
upar <- list(sf = gsDesign::sfLDOF, total_spend = 0.025, param = NULL, timing = NULL)
# Lower bound
lower <- gsdmvn::gs_spending_bound
lpar <- list(sf = gsDesign::sfHSD, total_spend = 0.1, param = 0, timing = NULL)
# weight function in WLR
wgt00 <- function(x, arm0, arm1){
                 gsdmvn:::wlr_weight_fh(x, arm0, arm1, rho = 0, gamma = 0)}
wgt05 <- function(x, arm0, arm1){
                 gsdmvn:::wlr_weight_fh(x, arm0, arm1, rho = 0, gamma = .5)}
# test in COMBO
fh_test <- rbind(
      data.frame(rho = 0, gamma = 0, tau = -1, test = 1, Analysis = 1:3,analysisTimes = c(12, 24, 36)), 
      data.frame(rho = c(0, 0.5), gamma = 0.5, tau = -1, test = 2:3, Analysis = 3, analysisTimes = 36)
      )
```

# `gs_design_ahr`
The previous outputs of `gs_design_ahr` is 

```{r, eval=FALSE}
# AHR design 
x_ahr <- gs_design_ahr(
   enrollRates = enrollRates,
   failRates = failRates,
   # Information fraction
   IF = IF, 
   # Analysis times in months; first 2 will be ignored as IF will not be achieved
   analysisTimes = analysisTimes, 
   # Experimental / Control randomization ratio
   ratio = ratio, 
   # 1-sided Type I error
   alpha = alpha, 
   # Type II error (1 - power)
   beta = beta, 
   upper = upper,
   upar = upar,
   lower = lower,
   lpar = lpar
)

x_ahr$bounds %>% 
  gt::gt()
```

The updated outputs of `gs_design_ahr2` is
```{r}
x_ahr2 <- gs_design_ahr2(
   enrollRates = enrollRates,
   failRates = failRates,
   IF = IF, # Information fraction
   analysisTimes = analysisTimes, # Analysis times in months; first 2 will be ignored as IF will not be achieved
   ratio = ratio, # Experimental / Control randomization ratio
   alpha = alpha, # 1-sided Type I error
   beta = beta, # Type II error (1 - power)
   upper = upper,
   upar = upar,
   lower = lower,
   lpar = lpar
)
```

The outputs of `gs_design_ahr2()` is a list including

- `enrollRates`: the enrollment rates
- `failRates`: the failure rates
- `bounds`: the summary of the crossing probabilities under H0 \& H1
- `analysis`: the summary of each analysis

The direct output of `x_ahr2$bounds` is
```{r}
x_ahr2$bounds %>% 
  gt::gt()
```

The direct output of `x_ahr2$analysis` is 
```{r}
x_ahr2$analysis %>% 
   gt::gt()
```

There is a nicer summary of the bounds by using `summary_bound()` function.  
```{r}
summary_bound(x_ahr2, method = "AHR", gt_style = FALSE) %>% 
   gt::gt()
```

One can also use `gt` to get the the above R tibble into a gt table.
```{r}
summary_bound(x_ahr2, method = "AHR", gt_style = TRUE)
```

If one set `gt_style = TRUE`, then one can edit the title/subtitle/spanner/footnotes of the gt table by using the arguments in `summary_bound`. 
But one can also use functions in `gt` for cumstom design of table layout. 
```{r}
summary_bound(x_ahr2, method = "AHR", gt_style = TRUE,
              gt_title = "Summary of the Crossing Probability",
              gt_subtitle = "by Using gs_design_ahr",
              gt_colname_spanner = "Cumulative boundary crossing probability",
              gt_colname_spannersub = c("Alternate hypothesis", "Null hypothesis"),
              gt_footnote = list(content = c("approximate hazard ratio to cross bound.", 
                                             "gs_design_ahr is a function in gsdmvn.",
                                             "AHR is average hazard ratio; IF is information fraction."),
                                 location = c("~HR at bound",
                                              NA, 
                                              NA),
                                 attr = c("colname",
                                          "subtitle", 
                                          "analysis")))
```


# `gs_design_wlr`

The previous outputs of `gs_design_wlr` is 
```{r, eval=FALSE}
x_wlr <- gsdmvn::gs_design_wlr(
   enrollRates = enrollRates,
   failRates = failRates,
   weight = wgt05, 
   # Information fraction
   IF = NULL, 
   # Analysis times in months; first 2 will be ignored as IF will not be achieved
   analysisTimes = sort(unique(x_ahr$bounds$Time)), 
   # Experimental / Control randomization ratio
   ratio = ratio, 
   # 1-sided Type I error
   alpha = alpha, 
   # Type II error (1 - power)
   beta = beta,   
   upper = upper,
   upar = upar,
   lower = lower,
   lpar = lpar
)

x_wlr$bounds %>% 
  gt::gt()
```

The updated outputs of `gs_design_wlr2` is
```{r, eval=FALSE}
x_wlr2 <- gs_design_wlr2(
   enrollRates = enrollRates,
   failRates = failRates,
   weight = wgt05, 
   # Information fraction
   IF = NULL, 
   # Analysis times in months; first 2 will be ignored as IF will not be achieved
   analysisTimes = sort(unique(x_ahr$bounds$Time)), 
   # Experimental / Control randomization ratio
   ratio = ratio, 
   # 1-sided Type I error
   alpha = alpha, 
   # Type II error (1 - power)
   beta = beta,   
   upper = upper,
   upar = upar,
   lower = lower,
   lpar = lpar
)

x_wlr2$bounds %>% 
  gt::gt()
```

The bounds can be summarized as 
```{r, eval=FALSE}
summary_bound(x_wlr2, method = "WLR")
```

# `gs_design_combo`

The original output of `gs_design_combo` is 
```{r, eval=FALSE}
x_combo <- gsdmvn::gs_design_combo(
  ratio = 1, 
  alpha = 0.025, 
  beta = 0.2, 
  enrollRates = tibble::tibble(Stratum = "All", duration = 12, rate = 500/12),
  failRates = tibble::tibble(Stratum = "All", duration = c(4, 100), 
                             failRate = log(2) / 15, hr = c(1, .6), dropoutRate = .001), 
  fh_test = fh_test,
  binding = FALSE,
  upper = gsdmvn::gs_spending_combo,
  upar = list(sf = gsDesign::sfLDOF, total_spend = 0.025),
  lower = gsdmvn::gs_spending_combo,
  lpar = list(sf = gsDesign::sfLDOF, total_spend = 0.2))

x_combo %>% gt::gt()
```              

The output of updated `gs_design_combo2` is
```{r, eval=FALSE}
x_combo2 <- gsdmvn::gs_design_combo2(
  ratio = 1, 
  alpha = 0.025, 
  beta = 0.2, 
  enrollRates = tibble::tibble(Stratum = "All", duration = 12, rate = 500/12),
  failRates = tibble::tibble(Stratum = "All", duration = c(4, 100), 
                             failRate = log(2) / 15, hr = c(1, .6), dropoutRate = .001), 
  fh_test = fh_test,
  upper = gsdmvn::gs_spending_combo,
  upar = list(sf = gsDesign::sfLDOF, total_spend = 0.025),
  lower = gsdmvn::gs_spending_combo,
  lpar = list(sf = gsDesign::sfLDOF, total_spend = 0.2))
```

```{r, eval=FALSE}
x_combo2 %>% gt::gt()
```

The bounds can be summarized as 
```{r, eval=FALSE}
summary_bound(x_combo2, method = "COMBO")
```


# `gs_power_npe`

The previous outputs of `gs_power_npe` is 
```{r, eval=FALSE}
gsdmvn::gs_power_npe(
  theta = c(.1, .2, .3), 
  info = (1:3) * 40, info0 = (1:3) * 40,
  upper = gs_b, upar = gsDesign::gsDesign(k = 3, sfu = gsDesign::sfLDOF)$upper$bound,
  lower = gs_b, lpar = c(-1, 0, 0)) %>% 
  gt::gt()
```

The updated outputs of `gs_power_npe2` is
```{r, eval=FALSE}
gs_power_npe2(
  theta = c(.1, .2, .3), 
  info = (1:3) * 40, info0 = (1:3) * 40,
  upper = gs_b, upar = gsDesign::gsDesign(k = 3, sfu = gsDesign::sfLDOF)$upper$bound,
  lower = gs_b, lpar = c(-1, 0, 0)) %>% 
  gt::gt()
```

# `gs_design_npe`
The previous outputs of `gs_design_npe` is 
```{r, eval=FALSE}
gsdmvn::gs_design_npe(
  theta = c(.1, .2, .3), 
  info = (1:3) * 80, info0 = (1:3) * 80, info1 = (1:3) * 80,
  upper = gs_b, upar = gsDesign::gsDesign(k = 3,sfu = gsDesign::sfLDOF)$upper$bound,
  lower = gs_b, lpar = c(-1, 0, 0)) %>% 
  gt::gt()
```

The updated outputs of `gs_design_npe2` is
```{r, eval=FALSE}
gs_design_npe2(
  theta = c(.1, .2, .3), 
  info = (1:3) * 80, info0 = (1:3) * 80, info1 = (1:3) * 80,
  upper = gs_b, upar = gsDesign::gsDesign(k = 3,sfu = gsDesign::sfLDOF)$upper$bound,
  lower = gs_b, lpar = c(-1, 0, 0))  %>% 
  gt::gt()
```

