---
title: "User stories for spending bounds"
output:
  rmarkdown::html_vignette:
    css: "custom.css"
    df_print: paged
    toc: true
    toc_depth: 2
    number_sections: true
bibliography: gsDesign.bib
vignette: |
  %\VignetteIndexEntry{User stories for spending bounds}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(tibble)
library(gt)
devtools::load_all()
```

# Overview

We assume a 2-arm group sequential design with an experimental treatment group and a control group.
Spending is used to specify boundary crossing probabilities for group sequential designs.
The simplest and most common application is to define upper (efficacy) boundaries that control $\alpha$-spending (Type I error) for a design. 
Lower bounds may be defined to be 

1) symmetric with upper efficacy bounds to do 2-sided testing,
2) asymmetric with
    a) futility bounds defined by $\beta$-spending to control Type II error,
    b) asymmetric lower bounds for harm, controlling probability that the experimental arm yields worse efficacy than control.
    
This vignette discusses considerations for distributional assumptions used to compute these bounds.
 
# References
